Parsing Postfix log files

### 2007-11-28 doc/logparser.tex
Split the flow chart into two, and tweaked the parts to be clearer. Fixed up all the references to the flow chart in the algorithm section.

XXX MAYBE REMOVE THE NEWPAGE BELOW WHEN I'VE SPLIT THE FLOW CHART\@.

\newpage
\subsection{Flow chart}

@@

Figure~\ref{flow chart image} shows the paths the data representing a
mail/connection can take through the parser algorithm.  The flow chart
covers the most common paths only; there are additional, uncommon paths
which are excluded for the sake of clarity, but are described in
section~\refwithpage{additional complications}.

XXX SPLIT THIS IMAGE INTO TWO\@; IT'S UNREADABLE NOW, IT'S SO BIG\@.

% This is very likely to move around as the flow chart is as big as
% possible, so if any preceding text changes the flow chart may end up on
% the next page, which may not be a bad thing, maybe.  What would be a bad
% thing is the \clearpage seems to cause a blank page sometimes, if the
% image is just slightly too big.
\showgraph{build/logparser-flow-chart.ps}{Parser flow chart}{flow chart
image}
\clearpage

@@

    \item Postfix will log the sender and recipient addresses separately
        (action: save\_by\_queueid; flowchart:~10).

    \item Sometimes mail is re-injected and the child mail needs to be
        tracked by the parent mail (action: track; flowchart:~11) ---
        section~\refwithpage{tracking re-injected mail} discusses this in
        detail.

    \item Eventually the mail will be delivered, bounced, or deleted by the
        administrator (action: commit; flowchart:~13).  This is the last
        log line for this particular mail (though it may be indirectly
        referred to if it was re-injected).  If it is neither parent nor
        child of re-injection the data is cleaned up and entered in the
        database (flowchart:~15), then deleted from the state tables
        (flowchart:~16).  Re-injected mails are described in
        section~\refwithpage{tracking re-injected mail}.

@@

    \item Local delivery re-injects the mail and logs a relayed delivery
        rather than delivering directly to a mailbox or program as it
        usually would (action: track; flowchart:~11).\footnote{Relayed
        delivery is performed by the \SMTP{} client; local delivery means
        local to the server, i.e.\ an address the server is final
        destination for.} In this case the mail may already have been
        created (described above) and the unknown origin flag will be
        cleared; if not a new data structure will be created.  In both
        cases the new mail is marked as a child of the parent.  The log
        line in question is:

@@

    \item If the mail is of unknown origin it is assumed to be a child mail
        whose parent hasn't yet been identified (action: commit;
        flowchart:~17).  Mark the mail as ready for entry in the database
        (flowchart:~18), and wait for the parent to deal with it
        (flowchart:~19).  The mail should not have subsequent log entries;
        only its parent should refer to it.

    \item If the mail is a child mail then it has already been tracked
        (action: commit; flowchart:~20): the data is cleaned up, the child
        is entered in the database (flowchart:~21), then deleted from the
        state tables (flowchart:~22).  The child mail will be removed from
        the parent mail's list of children (flowchart:~23); if this is the
        last child and the parent has also been entered in the database the
        parent will be deleted from the state tables.

    \item The last alternative is that the mail is a parent mail (action:
        commit; flowchart:~24).  Regardless of the state of its children
        its data is cleaned up and entered in the database (flowchart:~25).
        The parent may have children which are waiting to be entered in the
        database (flowchart:~26); each of those children's data is cleaned
        up and entered in the database, then deleted from the state tables.
        The parent may also have outstanding children which are not yet
        delivered, in which case (flowchart:~27) the parent must wait for
        those children to be finished with.  As soon as the last child is
        deleted from the state tables the parent will also be finished with
        (flowchart:~28), and deleted from the state tables.

@@

XXX CAN I PUT \% ON THE Y AXIS\@?

>>>>

\subsection{Flow chart}

@@

Figures~\refwithpage{flow chart image part 1} and~\refwithpage{flow chart
image part 2} shows the paths the data representing a mail/connection can
take through the parser algorithm.  The flow chart covers the most common
paths only; there are additional, uncommon paths which are excluded for the
sake of clarity, but are described in section~\refwithpage{additional
complications}.

% This is very likely to move around as the flow chart is as big as
% possible, so if any preceding text changes the flow chart may end up on
% the next page, which may not be a bad thing, maybe.  What would be a bad
% thing is the \clearpage seems to cause a blank page sometimes, if the
% image is just slightly too big.

\showgraph{build/logparser-flow-chart-part-1.ps}{Parser flow chart part
1}{flow chart image part 1}

\showgraph{build/logparser-flow-chart-part-2.ps}{Parser flow chart part
2}{flow chart image part 2}

@@

    \item Postfix will log the sender and recipient addresses separately
        (action: save\_by\_queueid; flowchart:~9).

    \item Sometimes mail is re-injected and the child mail needs to be
        tracked by the parent mail (action: track; flowchart:~10) ---
        section~\refwithpage{tracking re-injected mail} discusses this in
        detail.

    \item Eventually the mail will be delivered, bounced, or deleted by the
        administrator (action: commit; flowchart:~12).  This is the last
        log line for this particular mail (though it may be indirectly
        referred to if it was re-injected).  If it is neither parent nor
        child of re-injection the data is cleaned up and entered in the
        database (flowchart:~14), then deleted from the state tables.
        Re-injected mails are described in section~\refwithpage{tracking
        re-injected mail}.

@@

    \item Local delivery re-injects the mail and logs a relayed delivery
        rather than delivering directly to a mailbox or program as it
        usually would (action: track; flowchart:~10).\footnote{Relayed
        delivery is performed by the \SMTP{} client; local delivery means
        local to the server, i.e.\ an address the server is final
        destination for.} In this case the mail may already have been
        created (described above) and the unknown origin flag will be
        cleared; if not a new data structure will be created.  In both
        cases the new mail is marked as a child of the parent.  The log
        line in question is:

@@

    \item If the mail is of unknown origin it is assumed to be a child mail
        whose parent hasn't yet been identified (action: commit;
        flowchart:~15).  Mark the mail as ready for entry in the database
        (flowchart:~16), and wait for the parent to deal with it
        (flowchart:~17).  The mail should not have subsequent log entries;
        only its parent should refer to it.

    \item If the mail is a child mail then it has already been tracked
        (action: commit; flowchart:~18): the data is cleaned up, the child
        is entered in the database (flowchart:~19), then deleted from the
        state tables.  The child mail will be removed from the parent
        mail's list of children (flowchart:~20); if this is the last child
        and the parent has also been entered in the database the parent
        will be deleted from the state tables.

    \item The last alternative is that the mail is a parent mail (action:
        commit; flowchart:~21).  Regardless of the state of its children
        its data is cleaned up and entered in the database (flowchart:~22).
        The parent may have children which are waiting to be entered in the
        database (flowchart:~23); each of those children's data is cleaned
        up and entered in the database, then deleted from the state tables.
        The parent may also have outstanding children which are not yet
        delivered, in which case (flowchart:~24) the parent must wait for
        those children to be finished with.  As soon as the last child is
        deleted from the state tables the parent will also be finished with
        (flowchart:~25), and deleted from the state tables.

@@
